From Reversible Programs to \\ Univalent Universes and Back

### 2021-06-27 Pi+/paper/finite.tex
Proofs of crit pairs
Progress
Update finite.tex

When swapping two distinct pairs of elements (i.e. the indices of transpositions
differ by at least 1), it does not matter in which order the swapping happens,
that is, we can slide the wires freely.

@@

Unfortunately, in the system defined, this is not true for all critical pairs. One example is the following.

@@

As element $\el{0}$ is mapped to $\el{1}$, we remember that the last element of
result tuple is $1$. The appropriate elements on both sides are removed,
forming, respectively, $\FinExcept{4}[0]$ and $\FinExcept{4}[1]$. These are then
both transformed to $\Fin[3]$ using
\cref{prop:fin-finexcept}.~\ref{prop:fin-finexcept-2}, as shown below.
\[
    \begin{bmatrix}
        0 \\
        1 \\
        2
    \end{bmatrix}
    \equiv
    \begin{bmatrix}
        1 \\
        2 \\
        3
    \end{bmatrix}
    \mapsto
    \begin{bmatrix}
        0 \\
        2 \\
        3
    \end{bmatrix}
    \equiv
    \begin{bmatrix}
        0 \\
        1 \\
        2
    \end{bmatrix}
\]

>>>>

When swapping two distinct pairs of elements (i.e. when the indices of two transpositions
differ by at least 1), it does not matter in which order the swapping happens,
that is, we can slide the wires freely.

@@

\begin{proof}
\begin{array}{lcr}
% https://q.uiver.app/?q=WzAsNixbMiwwLCJcXHRhdV8yXFx0YXVfMVxcdGF1XzJcXHRhdV8xXFx0YXVfMiJdLFswLDEsIlxcdGF1XzFcXHRhdV8yXFx0YXVfMVxcdGF1XzFcXHRhdV8yIl0sWzQsMSwiXFx0YXVfMlxcdGF1XzFcXHRhdV8xXFx0YXVfMlxcdGF1XzEiXSxbMCwyLCJcXHRhdV8xXFx0YXVfMlxcdGF1XzIiXSxbMiwzLCJcXHRhdV8xIl0sWzQsMiwiXFx0YXVfMlxcdGF1XzJcXHRhdV8xIl0sWzAsMSwiYnJhaWQiLDIseyJzdHlsZSI6eyJib2R5Ijp7Im5hbWUiOiJzcXVpZ2dseSJ9fX1dLFswLDIsImJyYWlkIiwwLHsic3R5bGUiOnsiYm9keSI6eyJuYW1lIjoic3F1aWdnbHkifX19XSxbMSwzLCJjYW5jZWwiLDIseyJzdHlsZSI6eyJib2R5Ijp7Im5hbWUiOiJzcXVpZ2dseSJ9fX1dLFszLDQsImNhbmNlbCIsMix7InN0eWxlIjp7ImJvZHkiOnsibmFtZSI6InNxdWlnZ2x5In19fV0sWzIsNSwiY2FuY2VsIiwwLHsic3R5bGUiOnsiYm9keSI6eyJuYW1lIjoic3F1aWdnbHkifX19XSxbNSw0LCJjYW5jZWwiLDAseyJzdHlsZSI6eyJib2R5Ijp7Im5hbWUiOiJzcXVpZ2dseSJ9fX1dXQ==
\[\begin{tikzcd}
	&& {\tau_2\tau_1\tau_2\tau_1\tau_2} \\
	{\tau_1\tau_2\tau_1\tau_1\tau_2} &&&& {\tau_2\tau_1\tau_1\tau_2\tau_1} \\
	{\tau_1\tau_2\tau_2} &&&& {\tau_2\tau_2\tau_1} \\
	&& {\tau_1}
	\arrow["braid"', squiggly, from=1-3, to=2-1]
	\arrow["braid", squiggly, from=1-3, to=2-5]
	\arrow["cancel"', squiggly, from=2-1, to=3-1]
	\arrow["cancel"', squiggly, from=3-1, to=4-3]
	\arrow["cancel", squiggly, from=2-5, to=3-5]
	\arrow["cancel", squiggly, from=3-5, to=4-3]
\end{tikzcd}\]
&
\text{and}
&
% https://q.uiver.app/?q=WzAsNSxbMiwwLCJcXHRhdV8yXFx0YXVfMVxcdGF1XzJcXHRhdV8yIl0sWzAsMSwiXFx0YXVfMVxcdGF1XzJcXHRhdV8xXFx0YXVfMiJdLFsyLDQsIlxcdGF1XzJcXHRhdV8xIl0sWzAsMiwiXFx0YXVfMVxcdGF1XzJcXHRhdV8xXFx0YXVfMiJdLFsxLDMsIlxcdGF1XzFcXHRhdV8xXFx0YXVfMlxcdGF1XzEiXSxbMCwxLCJicmFpZCIsMix7InN0eWxlIjp7ImJvZHkiOnsibmFtZSI6InNxdWlnZ2x5In19fV0sWzAsMiwiY2FuY2VsIiwwLHsic3R5bGUiOnsiYm9keSI6eyJuYW1lIjoic3F1aWdnbHkifX19XSxbMSwzLCJicmFpZCIsMix7InN0eWxlIjp7ImJvZHkiOnsibmFtZSI6InNxdWlnZ2x5In19fV0sWzMsNCwiYnJhaWQiLDIseyJzdHlsZSI6eyJib2R5Ijp7Im5hbWUiOiJzcXVpZ2dseSJ9fX1dLFs0LDIsImNhbmNlbCIsMix7InN0eWxlIjp7ImJvZHkiOnsibmFtZSI6InNxdWlnZ2x5In19fV1d
\[\begin{tikzcd}
	&& {\tau_2\tau_1\tau_2\tau_2} \\
	{\tau_1\tau_2\tau_1\tau_2} \\
	{\tau_1\tau_2\tau_1\tau_2} \\
	& {\tau_1\tau_1\tau_2\tau_1} \\
	&& {\tau_2\tau_1}
	\arrow["braid"', squiggly, from=1-3, to=2-1]
	\arrow["cancel", squiggly, from=1-3, to=5-3]
	\arrow["braid"', squiggly, from=2-1, to=3-1]
	\arrow["braid"', squiggly, from=3-1, to=4-2]
	\arrow["cancel"', squiggly, from=4-2, to=5-3]
\end{tikzcd}\]
\end{array}
\end{proof}

Unfortunately, in the system defined, this is not true for all critical pairs.
One example is the following, where both endpoints are normal with respect to
$\cox*$ relation.

@@

We remember that the last element of result tuple is $k - 1 = 2$. The appropriate elements on both sides are removed,
forming, respectively, $\FinExcept{4}[0]$ and $\FinExcept{4}[1]$. These are then
both transformed to $\Fin[3]$ using
\cref{prop:fin-finexcept}.~\ref{prop:fin-finexcept-2}, as shown below.

\[
    \begin{bmatrix}
        0 \\
        1 \\
        2
    \end{bmatrix}
    \equiv
    \begin{bmatrix}
        1 \\
        2 \\
        3
    \end{bmatrix}
    \mapsto
    \begin{bmatrix}
        0 \\
        2 \\
        3
    \end{bmatrix}
    \equiv
    \begin{bmatrix}
        0 \\
        1 \\
        2
    \end{bmatrix}
\]
